The Companion Mass Distribution of extremely low-mass white dwarfs

### 2014-10-15 documents/letter/ms.tex
Added product equation

We construct a statistical model to derive constraints on a parametric model for the distribution of ELM WD companion masses, $p(M_2 \given \bs{\theta})$.\footnote{In what follows, vectors or sets of parameters or quantites are represented by bold symbols.} For each system, we assume we are given $K$, $T$, and $M_1$ and therefore know the value of the mass function, $\mf$ (\eqn~\ref{eq:massfunc}). We would like to derive posterior constraints on the parameters, $\bs{\theta}$, that describe the distribution of companion masses, $p(M_2\given \bs{\theta})$, given all of the observed mass functions, $\bs{m}_f$, by deconvolving the distribution of mass functions from the unobserved inclination angle, $i$. Using Bayes' rule,
\begin{equation}
    p(\bs{\theta} \given \mf) = \frac{1}{\mathcal{Z}}~p(\mf \given \bs{\theta})~p(\bs{\theta}).
\end{equation}
where $p(\mf \given \bs{\theta})$ is the likelihood, $p(\bs{\theta})$ is the prior on parameters $\bs{\theta}$, and the evidence integral, $\mathcal{Z}$, is a constant that depends only on the data. The marginal likelihood, $p(\mf \given \bs{\theta})$, involves integrals over the latent variables $i$ and $M_2$,
\begin{align}
    p(\mf \given \bs{\theta}) &= \int_0^{\pi/2} di \int_{-\infty}^\infty dM_2  \nonumber \\
      & \qquad {} \times p(\mf \given M_1, M_2, i)~p(M_2 \given \bs{\theta})~p(i).
\end{align}
We neglect observational uncertainties in $\mf$ and $M_1$,\footnote{This is justified because the relative uncertainties in these quantities are small, $\sigma_x / x \sim 0.05-0.1$.} and assume the inclination angles are isotropically distributed
\begin{align}
	p(\mf \given M_1, M_2, i) &= \delta \left[\mf - f(M_1, M_2, i) \right]\\
	f(M_1, M_2, i) &= \frac{(M_2 \sin i)^3}{(M_1 + M_2)^2}\\
	p(i) &= \sin i.
\end{align}
For now, we keep this general by not yet specifying a parametric form for the companion mass distribution, $p(M_2 \given \bs{\theta})$. With the above assumptions, the marginal likelihood integral is
\begin{align}
%    p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta}) \int_0^{\pi/2} di \nonumber \\
%    & \qquad {} \times \delta \left[\mf - f(M_1, M_2, i)\right]~\sin i\\
    p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2 ~p(M_2 \given \bs{\theta}) \int_0^{\pi/2} di  \nonumber \\
    & \qquad {} \times ~\sin i ~ \delta \left[g(M_1,M_2,i) \right]\label{eq:delta}
\end{align}
where
\begin{equation}
	g(M_1,M_2,i) = \frac{M_2^3}{(M_1+M_2)^2}\sin^3 i - \mf.
\end{equation}
The inner integral (over $i$) has the form
\begin{equation}
    \int dx~F(x)~\delta \left[ G(x) \right] = \sum_j \frac{F(x^*_j)}{|G'(x^*_j)|}
\end{equation}
where the sum is over the roots, $x^*_j$, of the function $G(x)$. The root, $i^*$, and derivative of the argument of the delta function in \eqn\ref{eq:delta} are 
\begin{align}
%	\sin i^* &= \alpha\\
%	\alpha &= \frac{(\mf(M_1+M_2)^2)^{1/3}}{M_2}\\
	\sin i^* &= \frac{ \left[\mf(M_1+M_2)^2 \right]^{1/3}}{M_2}\\
%	\frac{\partial g}{\partial i} &= \frac{3M_2^3}{(M_1+M_2)^2}\sin^2 i \cos i\\
%	\frac{\partial g}{\partial i}\bigg\rvert_{i^*} &= \frac{3M_2^3}{(M_1+M_2)^2} \alpha^2 \sqrt{1 - \alpha^2}
	\frac{\partial g}{\partial i}\bigg\rvert_{i^*} &= \frac{3M_2^3}{(M_1+M_2)^2} \sin^2 i^* \sqrt{1 - \sin^2 i^*}
\end{align}
With this in mind, we may write the marginal likelihood as
\begin{align}
	p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta})~\sin i^* \left(\frac{\partial g}{\partial i}\bigg\rvert_{i^*}\right)^{-1}\\
%	&= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta})~\alpha\left[\frac{3M_2^3}{(M_1+M_2)^2} \alpha^2 \sqrt{1 - \alpha^2}\right]^{-1}\\
	&= \int_{M_{2,{\rm min}}}^\infty dM_2~p(M_2 \given \bs{\theta})~h(M_2, \mf, M_1) \label{eq:fullm2}
\end{align}

>>>>

We construct a statistical model to derive constraints on a parametric model for the distribution of ELM WD companion masses, $p(M_2 \given \bs{\theta})$.\footnote{In what follows, vectors or sets of parameters or quantities are represented by bold symbols.} For each system, we assume we are given $K$, $T$, and $M_1$ and therefore know the value of the mass function, $\mf$ (\eqn~\ref{eq:massfunc}). We would like to derive posterior constraints on the parameters, $\bs{\theta}$, that describe the distribution of companion masses, $p(M_2\given \bs{\theta})$, given all of the observed mass functions, $\bs{m}_f$, by deconvolving the distribution of mass functions from the unobserved inclination angle, $i$. Using Bayes' rule,
\begin{equation}
    p(\bs{\theta} \given \bs{\mf}) = \frac{1}{\mathcal{Z}}~p(\bs{\mf} \given \bs{\theta})~p(\bs{\theta}).
\end{equation}
where $p(\bs{\mf} \given \bs{\theta})$ is the likelihood, $p(\bs{\theta})$ is the prior on parameters $\bs{\theta}$, and the evidence integral, $\mathcal{Z}$, is a constant that depends only on the data. The likelihood, $p(\bs{\mf} \given \bs{\theta})$, can be split into a product over the likelihoods of individual systems:

\begin{equation}
p(\bs{\mf} \given \bs{\theta}) = \prod_j p(\mf \given \bs{\theta})
\end{equation}

where the product is over all each of the $j$ systems.  The marginal likelihood involves integrals over the latent variables $i$ and $M_2$,
\begin{align}
    p(\mf \given \bs{\theta}) &= \int_0^{\pi/2} di \int_{-\infty}^\infty dM_2  \nonumber \\
      & \qquad {} \times p(\mf \given M_1, M_2, i)~p(M_2 \given \bs{\theta})~p(i).
\end{align}
We neglect observational uncertainties in $\mf$ and $M_1$,\footnote{This is justified because the relative uncertainties in these quantities are small, $\sigma_x / x \sim 0.05-0.1$.} and assume the inclination angles are isotropically distributed
\begin{align}
	p(\mf \given M_1, M_2, i) &= \delta \left[\mf - f(M_1, M_2, i) \right]\\
	f(M_1, M_2, i) &= \frac{(M_2 \sin i)^3}{(M_1 + M_2)^2}\\
	p(i) &= \sin i.
\end{align}
For now, we keep this general by not yet specifying a parametric form for the companion mass distribution, $p(M_2 \given \bs{\theta})$. With the above assumptions, the marginal likelihood integral is
\begin{align}
%    p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta}) \int_0^{\pi/2} di \nonumber \\
%    & \qquad {} \times \delta \left[\mf - f(M_1, M_2, i)\right]~\sin i\\
    p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2 ~p(M_2 \given \bs{\theta}) \int_0^{\pi/2} di  \nonumber \\
    & \qquad {} \times ~\sin i ~ \delta \left[g(M_1,M_2,i) \right]\label{eq:delta}
\end{align}
where
\begin{equation}
	g(M_1,M_2,i) = \frac{M_2^3}{(M_1+M_2)^2}\sin^3 i - \mf.
\end{equation}
The inner integral (over $i$) has the form
\begin{equation}
    \int dx~F(x)~\delta \left[ G(x) \right] = \sum_j \frac{F(x^*_j)}{|G'(x^*_j)|}
\end{equation}
where the sum is over the roots, $x^*_j$, of the function $G(x)$. The root, $i^*$, and derivative of the argument of the delta function in \eqn\ref{eq:delta} are 
\begin{align}
%	\sin i^* &= \alpha\\
%	\alpha &= \frac{(\mf(M_1+M_2)^2)^{1/3}}{M_2}\\
	\sin i^* &= \frac{ \left[\mf(M_1+M_2)^2 \right]^{1/3}}{M_2}\\
%	\frac{\partial g}{\partial i} &= \frac{3M_2^3}{(M_1+M_2)^2}\sin^2 i \cos i\\
%	\frac{\partial g}{\partial i}\bigg\rvert_{i^*} &= \frac{3M_2^3}{(M_1+M_2)^2} \alpha^2 \sqrt{1 - \alpha^2}
	\frac{\partial g}{\partial i}\bigg\rvert_{i^*} &= \frac{3M_2^3}{(M_1+M_2)^2} \sin^2 i^* \sqrt{1 - \sin^2 i^*}
\end{align}
With this in mind, we may write the marginal likelihood as
\begin{align}
	p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta})~\sin i^* \left(\frac{\partial g}{\partial i}\bigg\rvert_{i^*}\right)^{-1}\\
%	&= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta})~\alpha\left[\frac{3M_2^3}{(M_1+M_2)^2} \alpha^2 \sqrt{1 - \alpha^2}\right]^{-1}\\
	&= \int_{M_{2,{\rm min}}}^\infty dM_2~p(M_2 \given \bs{\theta})~h(M_2, \mf, M_1) \label{eq:fullm2}
\end{align}
